Multi-level generalizations of collective spin dynamics and squeezing

### 2019-10-03 collective_sun/sun_notes.tex
added a comment on an amplitude-matching drive condition
added comment about simulating large-spin synthetic gauge fields

An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{-i\phi_j} c_{j,\up}^\dag c_{j,\dn}
    + e^{i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j$ indexes a lattice
site, and $\phi_j$ is a site-dependent phase.  The multi-level
analogue of \eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, under which
\begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1}.
\end{align}
The SU($n$) symmetry of collective interactions implies that the
interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
invariant under this gauge transformation, but the single-particle
Hamiltonian in \eqref{eq:H_lat} transforms as
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat_SOC}
\end{align}
where $\phi$ is the relative phase of the driving laser between
adjacent lattice sites.  In the basis of drive operators, this
Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_drive}
\end{align}
with effective inhomogeneous driving field strengths
\begin{align}
  B_L^{(\phi,q)} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}

@@

$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, under which

@@

energy).

>>>>

An driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{-i\phi_j} c_{j,\up}^\dag c_{j,\dn}
    + e^{i\phi_j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j$ indexes a lattice
site, and $\phi_j$ is a site-dependent phase.  The multi-level
analogue of \eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{-i\phi_j} T_{1,1}^{(j)}
    + e^{i\phi_j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi_j} D_{LM}^{(j)}
    + \sin\p{M\phi_j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes
\begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1}.
\end{align}
The SU($n$) symmetry of collective interactions implies that the
interaction Hamiltonian $H_{\t{int}}$ in \eqref{eq:H_int_drive} is
invariant under this gauge transformation, but the single-particle
Hamiltonian in \eqref{eq:H_lat} transforms as
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -J \sum_{q,\mu} \cos\p{qa-\mu\phi} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat_SOC}
\end{align}
where $\phi$ is the relative phase of the driving laser between
adjacent lattice sites.  In the basis of drive operators, this
Hamiltonian takes the form (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}^{(\phi)} = -J \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_drive}
\end{align}
with effective inhomogeneous driving field strengths
\begin{align}
  B_L^{(\phi,q)} = f_L\p{qa} \sqrt{\f{L+1/2}{n}}
  \sum_\mu \bk{I\mu,L,0;I\mu} f_L\p{\mu\phi},
  &&
  f_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases}.
\end{align}
Spin-orbit coupling with multi-level systems thus allows for the
simulation of synthetic gauge fields with large spin ($L>1$).  As we
will see in the following section, multi-laser drives can also
simulate large-spin synthetic fields with non-zero spin projection
($M\ne0$) onto a quantization axis.

@@

$c_{j\mu}^\dag \to e^{i\mu\phi_j} c_{j\mu}^\dag$, which takes

@@

energy).  If instead $\Omega_0=\pm\Omega_\pm$, then the amplitudes
$\Omega_{1,0}$, $\Omega_{2,0}$ and $\Omega_{1,1}$ vanish, leaving only
the drives $\D_{2,1}$ and $\D_{2,2}$.
